---
title: "SimPlus Vignette"
author: "Jeff Endelman"
date: "2024-06-23"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE,comment="##",
                      dpi=150)
opts_knit$set(root.dir="~/Software/SimPlus/vignettes/")
```

SimPlus was created to facilitate testing [COMA](https://github.com/jendelman/COMA) using [AlphaSimR](https://cran.r-project.org/web/packages/AlphaSimR/index.html).

The following code creates a founder population in mutation-drift equilibrium and sets the variance parameters for a trait with additive and dominance effects. 

```{r}
founders <- runMacs2(nInd=500,nChr=10,segSites=NULL,Ne=100,
                     mutRate=1e-8, ploidy=2,
                     histGen=NULL,histNe=NULL,genLen=1)
SP <- SimParam$new(founders)
SP$setTrackRec(TRUE)
SP$addTraitADG(nQtlPerChr = 100, mean = 0, var = 1, varDD = 0.5, 
               meanDD = 0.5, useVarA = FALSE)
SP$setVarE(varE=2)
pop <- newPop(founders, simParam=SP)
```

The following code runs 20 generations of mass selection at 1% inbreeding, using phenotypes for genetic merit. The population is saved after 5 generations to compare the response with genomic selection. 

```{r}
gp <- genParam(pop, simParam = SP)
total.gen <- 20
results <- data.frame(gen=1:total.gen, mu=numeric(total.gen), 
                      genicVar=numeric(total.gen), 
                      F.A=numeric(total.gen), n.parent=numeric(total.gen), 
                      n.mate=numeric(total.gen),
                      accuracy=numeric(total.gen))  
results$mu[1] <- as.numeric(gp$mu)
results$genicVar[1] <- as.numeric(gp$genicVarG) 

p.ref <- update_TP(pop, SP, geno.file="geno.csv.gz", 
                   pheno.file="pheno.csv", ped.file="ped.csv", 
                   gen.TP=0)
update_K(pop, SP, K.method="A", K.file="kinship.csv", ped.file="ped.csv")

for (gen in 1:total.gen) {
  print(sub("X",gen,"Generation X"))
  ans1 <- sim_OCS(pop, SP, n.progeny=500, dF=0.01, solver="MOSEK",
                  COMA.file="pheno.csv", K.file="kinship.csv")
  
  #optional, measure accuracy
  acc <- sim_accuracy(pop, SP, COMA.file="pheno.csv")
  results.GEBV$accuracy[gen] <- acc$ocs
  
  ans2 <- sim_mate(pop, SP, matings=ans1$om, total.progeny=500, min.progeny=5)
  results$n.parent[gen] <- ans2$n.parent
  results$n.mate[gen] <- nrow(ans2$matings)
  
  pop <- setPheno(ans2$progeny,simParam = SP) #AlphaSimR function
  gp <- genParam(pop, simParam = SP)
  results$mu[gen] <- as.numeric(gp$mu)
  results$genicVar[gen] <- as.numeric(gp$genicVarG) 
  
  update_TP(pop, SP, geno.file="geno.csv.gz", pheno.file="pheno.csv", 
            ped.file="ped.csv", gen.TP=3)
  results$F.A[gen] <- update_K(pop, SP, K.method="A", K.file="kinship.csv", 
                               ped.file="ped.csv")
  if (gen==5)
    sim_save("gen5.rda",pop,SP,"geno.csv.gz","pheno.csv","ped.csv",p.ref=p.ref)
}

kable(results,digits=3)
```

Even though there is no genomic selection in the above simulation, the function `update_TP` is still required to update the phenotype and pedigree files for OCS. The argument `gen.TP` controls the number of generations in the TP (including the candidates, which are selected after phenotyping). It is set at 3 to anticipate the simulation below, ensuring the full TP is available at the onset of GS in generation 5.

To simulate GS requires another function, sim_StageWise, to predict the marker effects using the [StageWise](https://github.com/jendelman/StageWise) package.

```{r}
gen5 <- sim_load("gen5.rda","geno.csv.gz","pheno.csv","ped.csv")
pop <- gen5$pop
SP <- gen5$SP
results.GEBV <- results

update_K(pop, SP, K.method="A", K.file="kinship.csv", ped.file="ped.csv")

for (gen in 6:total.gen) {
  print(sub("X",gen,"Generation X"))
  ans0 <- sim_StageWise("geno.csv.gz","pheno.csv",ploidy=2,
                        COMA.file="COMA.csv", gen.TP=3)
  
  ans1 <- sim_OCS(pop, SP, n.progeny=500, dF=0.01, solver="MOSEK",
                  COMA.file="COMA.csv", K.file="kinship.csv")
  
  acc <- sim_accuracy(pop, SP, COMA.file="COMA.csv", K.file="kinship.csv")
  results.GEBV$accuracy[gen] <- acc$ocs
  
  ans2 <- sim_mate(pop, SP, matings=ans1$om, total.progeny=500, min.progeny=5)
  results.GEBV$n.parent[gen] <- ans2$n.parent
  results.GEBV$n.mate[gen] <- nrow(ans2$matings)
  
  pop <- setPheno(ans2$progeny,simParam = SP) 
  gp <- genParam(pop, simParam = SP)
  results.GEBV$mu[gen] <- as.numeric(gp$mu)
  results.GEBV$genicVar[gen] <- as.numeric(gp$genicVarG) 
  
  update_TP(pop, SP, geno.file="geno.csv.gz", pheno.file="pheno.csv",
            ped.file="ped.csv", gen.TP=3)
  results.GEBV$F.A[gen] <- update_K(pop, SP, K.method="A", K.file="kinship.csv", 
                               ped.file="ped.csv")
}
```

Simulating OMA follows the same code as above but replaces `sim_OCS` with `sim_OMA`, which has an additional argument `max.parent`. To limit the size of the computational problem for OMA, OCS is first used to reduce the number of candidates to `max.parent`. 

```{r}
gen5 <- sim_load("gen5.rda","geno.csv.gz","pheno.csv","ped.csv")
pop <- gen5$pop
SP <- gen5$SP
results.OMA <- results

update_K(pop, SP, K.method="A", K.file="kinship.csv", ped.file="ped.csv")

for (gen in 6:total.gen) {
  print(sub("X",gen,"Generation X"))
  ans0 <- sim_StageWise("geno.csv.gz","pheno.csv",ploidy=2,
                        COMA.file="COMA.csv", gen.TP=3)
  
  ans1 <- sim_OMA(pop, SP, n.progeny=500, dF=0.01, solver="MOSEK",
                  COMA.file="COMA.csv", K.file="kinship.csv",
                  max.parent=200)
  
  acc <- sim_accuracy(pop, SP, COMA.file="COMA.csv", K.file="kinship.csv")
  results.OMA$accuracy[gen] <- acc$oma
  
  ans2 <- sim_mate(pop, SP, matings=ans1$om, total.progeny=500, min.progeny=5)
  results.OMA$n.parent[gen] <- ans2$n.parent
  results.OMA$n.mate[gen] <- nrow(ans2$matings)
  
  pop <- setPheno(ans2$progeny,simParam = SP) 
  gp <- genParam(pop, simParam = SP)
  results.OMA$mu[gen] <- as.numeric(gp$mu)
  results.OMA$genicVar[gen] <- as.numeric(gp$genicVarG) 
  
  update_TP(pop, SP, geno.file="geno.csv.gz", pheno.file="pheno.csv",
            ped.file="ped.csv", gen.TP=3)
  results.OMA$F.A[gen] <- update_K(pop, SP, K.method="A", K.file="kinship.csv", 
                               ped.file="ped.csv")
}

pdat <- rbind(data.frame(results, method="OCS-P"),
              data.frame(results.GEBV, method="OCS-GEBV"),
              data.frame(results.OMA, method="OMA"))
```

```{r,echo=F}
pdat <- system.file("vignette_data", "results.csv", package = "SimPlus")
```

```{r}
library(ggplot2)
library(ggpubr)
p1 <- ggplot(pdat,aes(x=gen,y=mu,colour=method)) + geom_line()
p2 <- ggplot(pdat,aes(x=gen,y=genicVar,colour=method)) + geom_line()
p3 <- ggplot(pdat,aes(x=gen,y=F.A,colour=method)) + geom_line()
p4 <- ggplot(pdat,aes(x=gen,y=accuracy,colour=method)) + geom_line()

ggarrange(p1,p2,p3,p4,nrow=2,ncol=2,common.legend=T)
```

